---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---

This template demonstrates many of the bells and whistles of the `reprex::reprex_document()` output format. The YAML sets many options to non-default values, such as using `#;-)` as the comment in front of output.

## Code style

Since `style` is `TRUE`, this difficult-to-read code (look at the `.Rmd` source file) will be restyled according to the Tidyverse style guide when it's rendered. Whitespace rationing is not in effect!

```{r}
x=1;y=2;z=x+y;z
```

## Quiet tidyverse

The tidyverse meta-package is quite chatty at startup, which can be very useful in exploratory, interactive work. It is often less useful in a reprex, so by default, we suppress this.

However, when `tidyverse_quiet` is `FALSE`, the rendered result will include a tidyverse startup message about package versions and function masking.

```{r, eval = requireNamespace("tidyverse", quietly = TRUE)}
library(tidyverse)
```

