---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---
## Data Cleaning
We decided to remove these columns: CASE_TYPE, FOLDER_TYPE, WORK, PERMIT_NUMBER, REFERENCEFILE, FOLDERRSN, LINK, APPLICANT_FIRST_NAME, APPLICANT_LAST_NAME, END_HOURS_MOV_VEH_MON_SAT, END_HOURS_MOV_VEH_SUN, PROPERTYRSN, START_HOURS_MOV_VEH_MON_SAT, START_HOURS_MOV_VEH_SUN, PROPX, PROPY, LOCATION, UPDATE_DATE, FOLDER_DESCRIPTION
```{r}
# install.packages("dplyr")  # this was for Nneoma, since it wasn't working for you
library(dplyr)

df <- read.csv("sound_ordinance_data.csv", na.strings = c("", "NA")) # ensured that blank strings are counting as NA

columns_to_remove <- c("CASE_TYPE", "FOLDER_TYPE", "WORK", "PERMIT_NUMBER", "REFERENCEFILE", "FOLDERRSN", "LINK", 
  "APPLICANT_FIRST_NAME", "APPLICANT_LAST_NAME","END_HOURS_MOV_VEH_MON_SAT", "END_HOURS_MOV_VEH_SUN", 
  "PROPERTYRSN", "START_HOURS_MOV_VEH_MON_SAT", "START_HOURS_MOV_VEH_SUN", "PROPX", "PROPY", "LOCATION", "UPDATE_DATE", "FOLDER_DESCRIPTION",
  "FOLDER_NAME", "EVENT_YEAR", "APPLICANT_ORGANIZATION", "START_DATE_3", "START_DATE_4")

df <- df %>% select(-all_of(columns_to_remove))


df <- df %>% rename(EVENT_YEAR = CALENDAR_YEAR_FOLDER_CREATED)
df <- df %>% filter(!is.na(COUNCIL_DISTRICT))
df <- df %>% filter(!is.na(CALENDAR_YEAR_FOLDER_CREATED))
df <- df %>% filter(!is.na(EVENT_YEAR))
df <- df %>% filter(!is.na(EXISTING_ZONING))
df <- df %>% filter(!is.na(ISSUED_BY))

# unique_counts <- sapply(df, function(col), length(unique(col)))
# print(unique_counts)

colSums(is.na(df))
df$START_DATE_2 <- !(df$START_DATE_2 == "" | is.na(df$START_DATE_2)) # flipped it to make majority of set false instead of true


df <- df %>% rename(MULTIPLE_START_DATES = START_DATE_2)

unique_values <- unique(df$EVENT_MONTH)
# unique_values <- unique(df$OMV_TUE_START_TIME)
unique_values

# table(df$OMV_TUE_END_TIME)
# table(df$ISSUED_BY)


```


```{r}
# GREG 

```


```{r}
# ESHI 

```


```{r}
# DANIEL 

```

Nneoma is cleaning the values START_HOURS_NOT_NEAR_HOME, END_HOURS_NOT_NEAR_HOME, STATUS, EXPIRATION_DATE	(YYYY-MM-DD),
IN_DATE (YYYY-MM-DD), ISSUE_DATE, ISSUED_BY (YYYY-MM-DD), START_DATE, and has_multiple_start_dates
```{r}
# NNEOMA 
#nrow(df)
#************** START_DATE has 1,458 missing values -- if event month is blank, then start_date will be blank
#df$START_DATE
unique(df$START_DATE)
df %>% filter(is.na(df$START_DATE)) %>% select(START_DATE, STATUS) #checking cols of null start_date

#ISSUE DATE -- everything is clean and no missing values
unique(df$ISSUE_DATE)
df %>% select(ISSUE_DATE)  #%>%  filter(is.na(ISSUE_DATE))#checking cols of null start_date

#IN DATE -- everything is clean and no missing values
unique(df$IN_DATE)
df %>% select(IN_DATE)#filter(is.na(IN_DATE))  %>%   #checking cols of null start_date
head(df$IN_DATE, 1000)

# *********EXPIRATION_DATE -- if start_date is NA, then ex_date is NA

unique(df$EXPIRATION_DATE)
df %>% filter(is.na(EXPIRATION_DATE) & STATUS=='Expired') %>% select(EXPIRATION_DATE, START_DATE, STATUS) #  %>% summarize(n())  #checking cols of null start_date
head(df$EXPIRATION_DATE, 1000)

# ISSUEd_BY -- everything is clean and no missing values
table(df$ISSUED_BY)
unique(df$ISSUED_BY)
df %>% filter(is.na(ISSUED_BY))# %>% select(EXPIRATION_DATE, START_DATE, STATUS) #  %>% summarize(n())  #checking cols of null start_date
head(df, 1)

# MULTIPLE_START_DATES
table(df$MULTIPLE_START_DATES)
unique(df$MULTIPLE_START_DATES)
df$MULTIPLE_START_DATES
df %>% filter(is.na(ISSUED_BY))# %>% select(EXPIRATION_DATE, START_DATE, STATUS) #  %>% summarize(n())  #checking cols of null start_date
head(df, 1)

head(df)
```


```{r}
# MAADHAV 

```